---
title: "Effect Sizes, CV, Q10, and Other Summary Stats"
author: "Sam Csik"
date: "6/25/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

All summary stats, Q~10~ values, etc. found in the text of the manuscript are calculated here.

###**Outline**

**Part 0** Import packages and data
**Part 1** Isolate averages by temperature for use in subsequent calculations
**Part 2** Calculate Q~10~ between temperatures
**Part 3** Calculate in-text values

###**Required Packages:**

    - tidyverse
    - here

###**Required Data:**

**(file path: data/metabolism/outputs)**

    - smr_summary_stats.csv
    - mmr_summary_stats.csv
    - aas_summary_stats.csv
    - fas_summary_stats.csv

**(file path: data/heart_rate/outputs)**

    - rhr_summary_stats.csv
    - processed_heart_rates.csv

**(file path: data/foraging/outputs)**

    - max_consump_summary_stats.csv
    - a_summary_stats.csv
    - h_summary_stats.csv
    - foraging_assay_data.csv
    
####**Part 0.** Import packages and data
```{r}
##############################
# load required packages
##############################

source(here::here("code", "00_libraries.R"))

##############################
# load data
##############################

smr_summary <- read_csv(here::here("data", "metabolism", "outputs", "smr_summary_stats.csv"))

mmr_summary <- read_csv(here::here("data", "metabolism", "outputs", "mmr_summary_stats.csv"))

aas_summary <- read_csv(here::here("data", "metabolism", "outputs", "aas_summary_stats.csv"))

fas_summary <- read_csv(here::here("data", "metabolism", "outputs", "fas_summary_stats.csv"))

rhr_summary <- read_csv(here::here("data", "heart_rate", "outputs", "rhr_summary_stats.csv"))

mc_summary <- read_csv(here::here("data", "foraging", "outputs", "max_consump_summary_stats.csv"))

a_summary <- read_csv(here::here("data", "foraging", "outputs", "a_summary_stats.csv"))

h_summary <- read_csv(here::here("data", "foraging", "outputs", "h_summary_stats.csv"))

all_hr <- read_csv(here::here("data", "heart_rate", "outputs", "processed_heart_rates.csv"))

foraging_assays <- read_csv(here::here("data", "foraging", "raw", "foraging_assay_data.csv"))
```

####**Part 1.** Isolate mean values for use in following calculations
```{r}
##############################
# smr averages
##############################

smr11_avg <- smr_summary[1,2]
smr16_avg <- smr_summary[2,2]
smr21_avg <- smr_summary[3,2]
smr26_avg <- smr_summary[4,2]

##############################
# mmr averages
##############################

mmr11_avg <- mmr_summary[1,2]
mmr16_avg <- mmr_summary[2,2]
mmr21_avg <- mmr_summary[3,2]
mmr26_avg <- mmr_summary[4,2]

##############################
# aas averages
##############################

aas11_avg <- aas_summary[1,2]
aas16_avg <- aas_summary[2,2]
aas21_avg <- aas_summary[3,2]
aas26_avg <- aas_summary[4,2]

##############################
# fas averages
##############################

fas11_avg <- fas_summary[1,2]
fas16_avg <- fas_summary[2,2]
fas21_avg <- fas_summary[3,2]
fas26_avg <- fas_summary[4,2]

##############################
# rhr averages
##############################

rhr11_avg <- rhr_summary[1,2]
rhr16_avg <- rhr_summary[2,2]
rhr21_avg <- rhr_summary[3,2]

##############################
# max consumption averages
##############################

mc11_avg <- mc_summary[1,2]
mc16_avg <- mc_summary[2,2]
mc21_avg <- mc_summary[3,2]
mc26_avg <- mc_summary[4,2]
```

####**Part 2.** Q~10~ calculations; Q~10~ = (R2/R1)^(10/(T2-T1)))
```{r}
##############################
# smr Q10
##############################

Q10_11to26 <- (smr26_avg/smr11_avg) ^(10/(26-11))
Q10_11to16 <- (smr16_avg/smr11_avg) ^(10/(16-11))
Q10_16to21 <- (smr21_avg/smr16_avg) ^(10/(21-16))
Q10_21to26 <- (smr26_avg/smr21_avg) ^(10/(26-21))

##############################
# mmr Q10
##############################

mmr_Q10_11to26 <- (mmr26_avg/mmr11_avg) ^(10/(26-11))
mmr_Q10_11to16 <- (mmr16_avg/mmr11_avg) ^(10/(16-11))
mmr_Q10_16to21 <- (mmr21_avg/mmr16_avg) ^(10/(21-16))
mmr_Q10_21to26 <- (mmr26_avg/mmr21_avg) ^(10/(26-21))

##############################
# rhr Q10
##############################

rhr_Q10_11to21 <- (rhr21_avg/rhr11_avg) ^(10/(21-11))
rhr_Q10_11to16 <- (rhr16_avg/rhr11_avg) ^(10/(16-11))
rhr_Q10_16to21 <- (rhr21_avg/rhr16_avg) ^(10/(21-16))

##############################
# max consumption Q10
##############################

mc_Q10_11to26 <- (mc26_avg/mc11_avg) ^(10/(26-11))
mc_Q10_11to16 <- (mc16_avg/mc11_avg) ^(10/(16-11))
mc_Q10_16to21 <- (mc21_avg/mc16_avg) ^(10/(21-16))
mc_Q10_21to26 <- (mc26_avg/mc21_avg) ^(10/(26-21))
```

####**Part 3.** Values reported in-text
```{r}
# "SMR increased linearly by approximately 110% for each 5°C increase in temperature"
((smr26_avg - smr11_avg)/smr11_avg * 100)/3

# "For example, inter-individual variation in SMR at 11°C was 1.7 times greater than that observed among individuals at 16°C""
smr11_cv <- smr_summary[1,5]
smr16_cv <- smr_summary[2,5]
smr16_cv/smr11_cv

# "On average, resting heart rate (fH) increased by approximately 88.8% for each 5°C increase in temperature "
((rhr21_avg - rhr11_avg)/rhr11_avg * 100)/2

# "Heart rates intermittently stopped, which corresponded to periods of apnea, however this phenomenon was most common in 11°C-acclimated lobsters where heart rates stopped in 7.7% of recordings, as compared to less than 0.3% of the recordings for 16 and 21°C-acclimated lobsters."
percent_hr_stop <- all_hr %>% 
  dplyr::group_by(temperature) %>% 
  dplyr::summarize(
    number_stop = sum(hr < 2),
    number_go = sum(hr > 2),
    perc_stopped  = number_stop/number_go
  )

# "Maximum consumption rates increased 9.6-fold from the lowest (11°C) to the highest (26°C) temperature"
mc26_avg/mc11_avg

# "Zero mussels were consumed in 50% of ad libitum feeding assays conducted on 11°C-acclimated lobsters"
foraging60_at11C <- foraging_assays %>% 
  filter(temp == "11", 
         Initial == "60") %>% # total ad lib assays at 11C = 18
  filter(Killed == "0") # total ad lib assays where 0 mussels were consumed at 11C = 9

# "Despite bottom temperatures that frequently approach and even surpass 11°C (Fig. 1) in Santa Barbara Channel kelp forests, lobsters had SMRs that were, on average, 69.6% lower than lobsters acclimated to the regional near maximum of 21°C."
(smr21_avg - smr11_avg)/smr21_avg * 100

# "Lobster movement and activity within holding tanks was minimal in comparison to lobsters in other treatments (personal observation) and individuals consumed no mussels in 54.4% of all feeding assays, regardless of mussel density"
foraging60 <- foraging_assays %>% 
  filter(temp == "11") %>%  # 90 total feeding assays at 11C
  filter(Killed == "0") # 49 feeding assays where 0 mussels were consumed at 11C

49/90

# "If we consider the approximate temperature range in the Santa Barbara Channel (11-21°C), we find that the thermal sensitivity of lobster maximum consumption rates is ~39-fold higher at the colder (i.e. 11-16°C) versus warmer (i.e. 16-21°C) end of this range"
mc_Q10_11to16/mc_Q10_16to21

# "For example, if we assume that the average maximum consumption rate at 21°C is the most that lobsters are capable of consuming in the Santa Barbara Channel region, lobsters are able to recover 91.4% of their maximum consumption capacity when temperatures increase from 11 to 16°C"
mc16_avg/mc21_avg

# "The increase in lobster SMR (Ea, SMR)) was 2-fold stronger than the increase in lobster maximum consumption rates (Ea, C)) across mid to high temperatures (16-26°C; Fig A3, Table A9), challenging the MTE prediction that temperature drives proportional increases in standard metabolic rate and predation (Brown et al. 2004)"
Ea_smr <- 0.7006
Ea_mc <- 0.3381
Ea_smr/Ea_mc
```

####**Part 4.** Create plot of CVs (Fig S2)
```{r}
##############################
# isolate CVs
##############################

SMR_cv <- smr_summary %>% mutate(type = rep("SMR")) %>% dplyr::rename(CV = SMR_cv) %>% select(temp, type, CV)

MMR_cv <- mmr_summary %>% mutate(type = rep("MMR")) %>% dplyr::rename(CV = MMR_cv) %>% select(temp, type, CV)

AS_cv <- aas_summary %>% mutate(type = rep("AS")) %>% dplyr::rename(CV = AS_cv) %>% select(temp, type, CV)

MC_cv <- mc_summary %>% mutate(type = rep("Max Consumption")) %>% dplyr::rename(CV = consump_cv) %>% select(temp, type, CV)

a_cv <- a_summary %>% mutate(type = rep("Attack Rate")) %>% dplyr::rename(CV = a_cv) %>% select(temp, type, CV)

h_cv <- h_summary %>% mutate(type = rep("Handling Time")) %>% dplyr::rename(CV = h_cv) %>% select(temp, type, CV)

##############################
# combine into single df
##############################

all_cv <- rbind(SMR_cv, MMR_cv, AS_cv, MC_cv, a_cv, h_cv) %>%
  dplyr::rename(Parameter = type) %>%
  mutate(Parameter = fct_relevel(Parameter, "SMR", "MMR", "AS", "Max Consumption", "Attack Rate", "Handling Time"))

##############################
# plot
##############################

CV_plot <- ggplot(all_cv, aes(x = temp, y = CV, fill = Parameter)) +
  geom_jitter(aes(shape = Parameter), size = 4, width = 0.2) +
  labs(x = "Temperature (°C)", y = "Coefficient of Variation") +
  theme_classic() + 
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fill = NA, size = 0.7),
        plot.caption = element_text(size = 10, hjust = 0))

# cowplot::save_plot(here::here("figures", "FigS2.pdf"), CV_plot, base_width = 5.5, base_height = 4)
```

