---
title: "Metabolic Trait Analysis"
author: "Sam Csik"
date: "6/23/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

We examine the relationship between lobster metabolic traits (SMR, MMR, AAS) and temperature by fitting linear and nonlinear models to the data and using AIC to determine the best fit model.

###**Outline**

**Part 0.** Import libraries and data
**Part 1.** (TABLE S5 GENERATED HERE) Compare linear and exponential model fits to SMR data
**Part 2.** (TABLE S6 GENERATED HERE) Compare linear and quadratic model fits to MMR data
**Part 3.** (TABLE S7 GENERATED HERE) Compare linear and quadratic model fits to AAS data

###**Packages Required:** 

    - sjPlot

###**Data Required:**

**(file path: data/metabolism)**

    - metabolic_traits.csv

####**Part 0.** Import packages and data   
```{r}
##############################
# import libraries
##############################

source(here::here("code", "0_libraries.R"))

##############################
# load data
##############################

metabolic_rates <- read_csv(here::here("data", "metabolism", "metabolic_traits.csv")) %>% 
  mutate(temp = as.numeric(as.character(temp)))

# epoc <- read_csv(here::here("data", "metabolism", "epoc.csv"))
# 
# epoc_sum <- read_csv(here::here("data", "metabolism", "outputs", "epoc_summary_stats.csv")) %>% 
#   mutate(temp = as.factor(temp))
```

####**Part 1.** Fit linear and exponential models to *SMR* data; create table for supplement
```{r}
##############################
# fit models
##############################

# linear model
smr_lm_mod <- lm(formula = SMR ~ temp, data = metabolic_rates)
summary(smr_lm_mod)

# exponential model
smr_exponential_mod <- lm(log(metabolic_rates$SMR) ~ log(metabolic_rates$temp))
summary(smr_exponential_mod)

##############################
# create table (manually edit aesthetics in Word)
##############################

tab_model(smr_lm_mod, smr_exponential_mod,
          dv.labels = c("Linear", "Exponential"),
          pred.labels = c(`(Intercept)` = "Intercept",
                          temp = "Temperature (°C)",
                          `all_mr_mod$temp [log]` = "log(Temperature(°C))"),
          string.pred = "Coefficient",
          string.est = "Estimate",
          show.se = TRUE,
          string.se = "SE",
          string.ci = "95% CI",
          show.aic = TRUE)
```

####**Part 2.** Fit linear and quadratic models to *MMR* data; create table for supplement
```{r}
##############################
# fit models
##############################

# linear model
mmr_lm_mod <- lm(formula = MMR ~ temp, data = metabolic_rates)
summary(mmr_lm_mod)

# quadratic model
temp2 <- metabolic_rates$temp^2 # temp squared
mmr_quadratic_mod <- lm(MMR ~ temp + temp2, data = metabolic_rates)
summary(mmr_quadratic_mod)

##############################
# create table (manually edit aesthetics in Word)
##############################

tab_model(mmr_lm_mod, mmr_quadratic_mod,
          dv.labels = c("Linear", "Quadratic"),
          pred.labels = c(`(Intercept)` = "Intercept",
                          temp = "Temperature (°C)",
                          `all_mr_mod$temp [log]` = "Temperature(°C)^2"),
          string.pred = "Coefficient",
          string.est = "Estimate",
          show.se = TRUE,
          string.se = "SE",
          string.ci = "95% CI",
          show.aic = TRUE)
```

####**Part 3.** Fit linear and quadratic models to *AAS* data; create table for supplement
```{r}
##############################
# fit models
##############################

# linear model
aas_lm_mod <- lm(formula = AAS ~ temp, data = metabolic_rates)
summary(aas_lm_mod)

# quadratic model
temp2 <- metabolic_rates$temp^2 # temp squared
aas_quadratic_mod <- lm(AAS ~ temp + temp2, data = metabolic_rates)
summary(aas_quadratic_mod)

##############################
# create table (manually edit aesthetics in Word)
##############################

tab_model(aas_lm_mod, aas_quadratic_mod,
          dv.labels = c("Linear", "Quadratic"),
          pred.labels = c(`(Intercept)` = "Intercept",
                          temp = "Temperature (°C)",
                          `all_mr_mod$temp [log]` = "Temperature(°C)^2"),
          string.pred = "Coefficient",
          string.est = "Estimate",
          show.se = TRUE,
          string.se = "SE",
          string.ci = "95% CI",
          show.aic = TRUE)
```

####**Part 4.** Plot hourly EPOC (0-5 hrs post MMR) for each temperature
```{r}
##############################
# wrangle EPOC data
##############################

# epoc_long <- epoc %>% 
#   gather(EPOC_1hr, EPOC_2hr, EPOC_3hr, EPOC_4hr, EPOC_5hr, key = hour, value = EPOC) %>%
#   mutate(temp = as.factor(temp)) %>%
#   mutate(
#     hour_labels = case_when(
#       hour == "EPOC_1hr" ~ "1",
#       hour == "EPOC_2hr" ~ "2",
#       hour == "EPOC_3hr" ~ "3",
#       hour == "EPOC_4hr" ~ "4",
#       hour == "EPOC_5hr" ~ "5"
#     )
#   )

##############################
# plot EPOC over time for each temp
##############################

# wrangle
# epoc_sum_new <- epoc_sum %>% 
#    mutate(
#     hour_labels = case_when(
#       hour == "EPOC_1hr" ~ "1",
#       hour == "EPOC_2hr" ~ "2",
#       hour == "EPOC_3hr" ~ "3",
#       hour == "EPOC_4hr" ~ "4",
#       hour == "EPOC_5hr" ~ "5"
#     )
#   )

# plot
# ggplot() + 
#   # geom_point(data = epoc_sum_new, aes(x = hour_labels, y = EPOC_avg, color = temp, shape = temp), size = 2) +
#   geom_line(data = epoc_sum_new, aes(x = hour_labels, y = EPOC_avg, color = temp)) +
#   scale_color_manual(values=c("lightslategray", "lightblue", "lightcoral", "indianred4")) +
#   labs(y = expression(atop("EPOC", paste("(",mg~O[2]~kg^-1,")"))),
#        x = expression(paste("Temperature (", degree, "C)"))) +
#   theme_classic() + 
#   theme(axis.text = element_text(color = "black", size = 12),
#         axis.title = element_text(size = 13),
#         panel.border = element_rect(colour = "black", fill = NA, size = 0.7),
#         plot.caption = element_text(size = 10, hjust = 0),
#         legend.position = "none")

```

