---
title: "Calculte Resting Heart Rate"
author: "Sam Csik"
date: "2/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

Calculate resting heart rate (f~H~) for each lobster as the 15th percentile of each lobster's heart rate measurements.
 
We also examine the relationship between lobster resting heart rate and temperature by fitting linear and quadratic models to the data and using AIC to determine the best fit model.

###**Outline**

**Part 0** Import packages and data
**Part 1** Calculate resting heart rate for each lobster
**Part 2** Calculate treatment-level summary stats (will eventually be used for plotting)
**Part 3** (TABLE S8 GENERATED HERE) Compare linear and quadratic model fits to resting heart rate data

###**Packages Required:** 

    - tidyverse
    - here
    - sjPlot
    
###**Data Required:**
    
**(file path: data/heart_rate/outputs)**

    - processed_heart_rates.csv

####**Part 0.** Import libraries and data
```{r}
##############################
# load required packages
##############################

source(here::here("code", "0_libraries.R"))

##############################
# load data
##############################

all_hr <- read_csv(here::here("data", "heart_rate", "outputs", "processed_heart_rates.csv")) %>% 
  select(date2, lobster_id, temperature, hr, QI) 
```

####**Part 1.** Calculate resting heart rate as the 15th percentile of each lobster's heart rate measurements (keeping consistent with SMR analyses)
```{r}
##############################
# 21C
##############################

L3 <- all_hr %>% filter(lobster_id == "L3")
L3q <- quantile(L3$hr, c(0.15)) # 61 

L4 <- all_hr %>% filter(lobster_id == "L4")
L4q <- quantile(L4$hr, c(0.15)) # 120

N14 <- all_hr %>% filter(lobster_id == "N14")
N14q <- quantile(N14$hr, c(0.15)) # 90

##############################
# 16C
##############################

GP05 <- all_hr %>% filter(lobster_id == "GP05") 
GP05q <- quantile(GP05$hr, c(0.15)) # 71

GP06 <- all_hr %>% filter(lobster_id == "GP06") 
GP06q <- quantile(GP06$hr, c(0.15)) # 75

IV17 <- all_hr %>% filter(lobster_id == "IV17") 
IV17q <- quantile(IV17$hr, c(0.15)) # 85

##############################
# 11C
##############################

IV11 <- all_hr %>% filter(lobster_id == "IV11") 
IV11q <- quantile(IV11$hr, c(0.15)) # 34

IV13 <- all_hr %>% filter(lobster_id == "IV13") 
IV13q <- quantile(IV13$hr, c(0.15)) # 41

IV14 <- all_hr %>% filter(lobster_id == "IV14") 
IV14q <- quantile(IV14$hr, c(0.15)) # 19.65

##############################
# make into df
##############################

resting_heart_rates <- data.frame("lob_id" = c("L3", "L4", "N14", "GP05", "GP06", "IV17", "IV11", "IV13", "IV14"),
                                  "temp" = c("21", "21", "21", "16", "16", "16", "11", "11", "11"),
                                  "rhr" = c(L3q, L4q, N14q, GP05q, GP06q, IV17q, IV11q, IV13q, IV14q))

# write.csv(resting_heart_rates, here::here("data", "heart_rate", "outputs", "resting_heart_rates.csv"), row.names = FALSE)
```

#####**Part 2.** Calculate summary stats (treatment-level)
```{r}
hr_summary <- resting_heart_rates %>% 
  dplyr::group_by(temp) %>% 
  dplyr::summarize(
    hr_avg = mean(rhr),
    hr_sd = sd(rhr),
    hr_se = hr_sd/sqrt(length(rhr)),
    hr_cv = hr_sd/hr_avg,
    hr_max = max(rhr),
    hr_min = min(rhr),
    error = qnorm(0.975) * sd(rhr)/sqrt(length(rhr)),
    upper_95ci = hr_avg + error,
    lower_95ci = hr_avg - error
    )

# write.csv(hr_summary, here::here("data", "heart_rate", "outputs", "rhr_summary_stats.csv"), row.names = FALSE)
```

####**Part 3.** Fit linear and quadratic models to *resting heart rate* data; create table for supplement
```{r}
##############################
# fit models
##############################

resting_heart_rates_numerictemps <- resting_heart_rates %>% mutate(temp = as.numeric(as.character(temp)))

##############################
# fit models
##############################

# linear model
hr_lm_mod <- lm(formula = rhr ~ temp, data = resting_heart_rates_numerictemps)
summary(hr_lm_mod)

# quadratic model
temp2 <- resting_heart_rates_numerictemps$temp^2 # temp squared
hr_quadratic_mod <- lm(rhr ~ temp + temp2, data = resting_heart_rates_numerictemps)
summary(hr_quadratic_mod)

##############################
# create table (manually edit aesthetics in Word)
##############################

tab_model(hr_lm_mod, hr_quadratic_mod,
          dv.labels = c("Linear Model", "Quadratic Model"),
          pred.labels = c(`(Intercept)` = "Intercept",
                          temp = "Temperature (°C)",
                          temp2 = "Temperature (°C)^2"),
          string.pred = "Coefficient",
          string.est = "Estimate",
          show.se = TRUE,
          string.se = "SE",
          string.ci = "95% CI",
          show.aic = TRUE)
```