---
title: "Foraging Assay Data Processing"
author: "Sam Csik"
date: "6/23/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

Maximum consumption rate (*C*) is calculated as the total number of mussels consumed in a 24 h feeding assay out of 60 mussels (we consider this to be an *ad lib* feeding trial, as no lobster ever ate all 60 mussels).

###**Outline**

**Part 0** Import packages and data
**Part 1** Calculate maximum consumption rates and summary stats for lobsters using *ad lib* feeding assay data (i.e. from trials with 60 mussels)

###**Packages Required:**

    - tidyverse
    - here

##**Data Required:**

**(file path: data/functional_response/raw)**

    - foraging_assay_data.csv

####**Part 0.** Import libraries and data
```{r}
##############################
# load required packages
##############################

source(here::here("code", "0_libraries.R"))

##############################
# load data
##############################

foraging_assays <- read_csv(here::here("data", "functional_response", "raw", "foraging_assay_data.csv")) 
```   

####**Part 1.** Calculate summary stats for max consuumption data
```{r}
##############################
# filter data for only trials with 60 mussels
##############################

foraging60 <- foraging_assays %>% 
  filter(Initial == "60")

##############################
# summary stats
##############################

consump_summary <- foraging60 %>% 
  mutate(temp = as.factor(temp)) %>% 
  dplyr::group_by(temp) %>% 
  dplyr::summarize(
    consump_avg = round(mean(Killed),3),
    consump_sd = round(sd(Killed),3),
    consump_se = consump_sd/sqrt(length(Killed)),
    consump_cv = consump_sd/consump_avg,
    consump_max = max(Killed),
    consump_min = min(Killed),
    error = qnorm(0.975) * sd(Killed)/sqrt(length(Killed)),
    upper_95ci = consump_avg + error,
    lower_95ci = consump_avg - error
  )

# write.csv(consump_summary, here::here("data", "functional_response", "outputs", "max_consump_summary_stats.csv"), row.names = FALSE)
```