---
title: "EPOC Analysis & Plotting"
author: "Sam Csik"
date: "7/8/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Summary**

###**Outline**

###**Packages Required:**

    - tidyverse
    - here

###**Data Required:**

**(file path: data/metabolism/outputs)**

    - epoc_summary_stats.csv
    
####**Part 0.** Import packages and data  
```{r}
##############################
# import libraries
##############################

source(here::here("code", "0_libraries.R"))

##############################
# load data
##############################

epoc <- read_csv(here::here("data", "metabolism", "epoc.csv")) %>% filter(spar == 0.2)

IV11 <- epoc %>% filter(ID == "IV11")
IV13 <- epoc %>% filter(ID == "IV13")
IV16 <- epoc %>% filter(ID == "IV16")
GP07 <- epoc %>% filter(ID == "GP07")

epoc_full_sum <- read_csv(here::here("data", "metabolism", "outputs", "epoc_full_summary_stats.csv"))

epoc_hourly_sum <- read_csv(here::here("data", "metabolism", "outputs", "epoc_hourly_summary_stats.csv")) %>% 
  mutate(temp = as.factor(temp))

time_to_epoc_sum <- read_csv(here::here("data", "metabolism", "outputs", "time_to_epoc_full_summary_stats.csv"))

epoc_at_MMR50_sum <- read_csv(here::here("data", "metabolism", "outputs", "epoc_at_MMR50_summary_stats.csv"))

time_to_MMR50_sum <- read_csv(here::here("data", "metabolism", "outputs", "time_to_MMR50_summary_stats.csv"))

MMR50_sum <- read_csv(here::here("data", "metabolism", "outputs", "MMR50_summary_stats.csv"))
```

####**Part 1.** Plot total EPOC for each temperature & run ANOVA
```{r}
##############################
# plot
##############################

full_EPOC_plot <- ggplot() +
  geom_jitter(data = epoc, aes(x = as.factor(temp), y = EPOC_full), color = "black", alpha = 0.2, width = 0.2, size = 3) + 
  # geom_point(data = IV11, aes(x = as.factor(temp), y = EPOC_full), color = "red", alpha = 0.5, size = 3) +
  # geom_point(data = IV13, aes(x = as.factor(temp), y = EPOC_full), color = "blue", alpha = 0.5, size = 3) +
  # geom_point(data = GP07, aes(x = as.factor(temp), y = EPOC_full), color = "green", alpha = 0.5, size = 3) +
  # geom_point(data = IV16, aes(x = as.factor(temp), y = EPOC_full), color = "purple", alpha = 0.5, size = 3) +
  geom_errorbar(data = epoc_full_sum, aes(x = as.factor(temp), ymin = epoc_full_avg - epoc_full_se, ymax = epoc_full_avg + epoc_full_se), width = 0.1, size = 1, color = "black") + # sub se for "error" to get 95% CI
  geom_point(data = epoc_full_sum, aes(x = as.factor(temp),  y = epoc_full_avg), size = 6, color = "black") +
  geom_point(data = epoc_full_sum, aes(x = as.factor(temp),  y = epoc_full_avg, color = as.factor(temp)), size = 4.5) +
  labs(y = expression("Cumulative EPOC (mg" ~ O[2] ~kg^-1*")"),
       x = "Temperature (°C)") +
  scale_color_manual(values = c("lightslategray", "lightblue", "lightcoral", "indianred4")) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fill = NA, size = 0.7),
        plot.caption = element_text(size = 10, hjust = 0),
        legend.position = "none")

##############################
# ANOVA
##############################

shapiro.test(epoc$EPOC_full) # p = 0.03425 not normal
ggqqplot(epoc$EPOC_full, ylab = "EPOC") 
leveneTest(EPOC_full ~ as.factor(temp), epoc) # p = 0.2559 equal variances

EPOC_anova <- aov(EPOC_full ~ as.factor(temp), data = epoc)
summary(EPOC_anova) 


```

####**Part 2.** Plot hourly EPOC for each temperature (REMOVE IV11??) (CURRENTLY SE)
```{r}
##############################
# plot
##############################

pd = position_dodge(0.4)

hourly_EPOC_plot <- ggplot(epoc_hourly_sum, aes(x = hour, y = mean, group = temp, color = temp)) +
  geom_line(position = pd, alpha = 0.6, size = 1.05) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, color = temp), width = 0.2, position = pd, alpha = 0.6) +
  geom_point(data = epoc_hourly_sum, aes(x = hour, y = mean, group = temp, color = temp, shape = temp, size = temp), position = pd) +
  scale_color_manual(values = c("lightslategray", "lightblue", "lightcoral", "indianred4"), name = "Temperature (°C)") +
  scale_shape_manual(values = c(15, 16, 17, 18), name = "Temperature (°C)", labels = c("11", "16", "21", "26")) +
  scale_size_manual(values = c(4, 4, 4, 5), name = "Temperature (°C)", labels = c("11", "16", "21", "26")) +
  labs(x = "Time post-exercise (h)",
       y = expression("Hourly EPOC (mg" ~ O[2] ~kg^-1*")")) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fill = NA, size = 0.7),
        plot.caption = element_text(size = 10, hjust = 0),
        legend.position = c(0.99, 0.99),
        legend.justification = c("right", "top"),
        legend.background = element_blank(),
        legend.box.just = "right",
        legend.title.align = 0.5)
```

####**Part 3.** Plot time to EPOC for each temperature (REMOVE IV11??)
```{r}
##############################
# plot
##############################

time_to_EPOC_plot <- ggplot() +
  geom_point(data = epoc, aes(x = as.factor(temp), y = end_EPOC_min), color = "black", alpha = 0.5, size = 3) + # width = 0.2, alpha = 0.2
  geom_point(data = IV11, aes(x = as.factor(temp), y = end_EPOC_min), color = "red", alpha = 0.5, size = 3) +
    geom_point(data = IV13, aes(x = as.factor(temp), y = end_EPOC_min), color = "blue", alpha = 0.5, size = 3) +
  geom_point(data = GP07, aes(x = as.factor(temp), y = end_EPOC_min), color = "green", alpha = 0.5, size = 3) +
  geom_point(data = IV16, aes(x = as.factor(temp), y = end_EPOC_min), color = "purple", alpha = 0.5, size = 3) +
  geom_errorbar(data = time_to_epoc_sum , aes(x = as.factor(temp), ymin = time_to_epoc_avg - time_to_epoc_se, ymax = time_to_epoc_avg + time_to_epoc_se), width = 0.1, size = 1, color = "black") + # sub se for "error" for 95% CI
  geom_point(data = time_to_epoc_sum , aes(x = as.factor(temp),  y = time_to_epoc_avg), size = 6, color = "black") +
  geom_point(data = time_to_epoc_sum, aes(x = as.factor(temp),  y = time_to_epoc_avg, color = as.factor(temp)), size = 4.5) +
  labs(y = "Time to EPOC (min)",
       x = "Temperature (°C)") +
  scale_color_manual(values = c("lightslategray", "lightblue", "lightcoral", "indianred4")) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fill = NA, size = 0.7),
        plot.caption = element_text(size = 10, hjust = 0),
        legend.position = "none")
```

####**Part 4.** Plot MMR~50~ for each temperature (REMOVE IV11??)
```{r}
EPOC_at_MMR50_plot <- ggplot() +
  geom_point(data = epoc, aes(x = as.factor(temp), y = EPOC_mmr), color = "black", alpha = 0.5, size = 3) + #  width = 0.2, alpha = 0.2,
  geom_point(data = IV11, aes(x = as.factor(temp), y = EPOC_mmr), color = "red", alpha = 0.5, size = 3) +
    geom_point(data = IV13, aes(x = as.factor(temp), y = EPOC_mmr), color = "blue", alpha = 0.5, size = 3) +
  geom_point(data = GP07, aes(x = as.factor(temp), y = EPOC_mmr), color = "green", alpha = 0.5, size = 3) +
  geom_point(data = IV16, aes(x = as.factor(temp), y = EPOC_mmr), color = "purple", alpha = 0.5, size = 3) +
  geom_errorbar(data = epoc_at_MMR50_sum , aes(x = as.factor(temp), ymin = epoc_at_MMR50_avg - epoc_at_MMR50_se, ymax = epoc_at_MMR50_avg + epoc_at_MMR50_se), width = 0.1, size = 1, color = "black") + # replace with error for 95% CI
  geom_point(data = epoc_at_MMR50_sum , aes(x = as.factor(temp),  y = epoc_at_MMR50_avg), size = 6, color = "black") +
  geom_point(data = epoc_at_MMR50_sum, aes(x = as.factor(temp),  y = epoc_at_MMR50_avg, color = as.factor(temp)), size = 4.5) +
  labs(y = expression("EPOC at" ~ MMR[50] ~"(mg" ~ O[2] ~ kg^-1*")"),
       x = "Temperature (°C)") +
  scale_color_manual(values = c("lightslategray", "lightblue", "lightcoral", "indianred4")) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fill = NA, size = 0.7),
        plot.caption = element_text(size = 10, hjust = 0),
        legend.position = "none")
```

####**Part 5.** Plot time to MMR~50~ for each temperature (REMOVE IV11??)
```{r}
##############################
# plot
##############################

time_to_MMR50_plot <- ggplot() +
  geom_point(data = epoc, aes(x = as.factor(temp), y = end_EPOC_mmr), color = "black", alpha = 0.5, size = 3) + # width = 0.2, alpha = 0.2,
  geom_point(data = IV11, aes(x = as.factor(temp), y = end_EPOC_mmr), color = "red", alpha = 0.5, size = 3) +
  geom_point(data = IV13, aes(x = as.factor(temp), y = end_EPOC_mmr), color = "blue", alpha = 0.5, size = 3) +
  geom_point(data = GP07, aes(x = as.factor(temp), y = end_EPOC_mmr), color = "green", alpha = 0.5, size = 3) +
  geom_point(data = IV16, aes(x = as.factor(temp), y = end_EPOC_mmr), color = "purple", alpha = 0.5, size = 3) +
  geom_errorbar(data = time_to_MMR50_sum , aes(x = as.factor(temp), ymin = time_to_MMR50_avg - error, ymax = time_to_MMR50_avg + error), width = 0.1, size = 1, color = "black") +
  geom_point(data = time_to_MMR50_sum , aes(x = as.factor(temp),  y = time_to_MMR50_avg), size = 6, color = "black") +
  geom_point(data = time_to_MMR50_sum, aes(x = as.factor(temp),  y = time_to_MMR50_avg, color = as.factor(temp)), size = 4.5) +
  labs(y = expression("Time to" ~ MMR[50] ~ "(min)"),
       x = "Temperature (°C)") +
  scale_color_manual(values = c("lightslategray", "lightblue", "lightcoral", "indianred4")) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fill = NA, size = 0.7),
        plot.caption = element_text(size = 10, hjust = 0),
        legend.position = "none")
```

####**Part 6.** Plot MMR~50~ for each temperature (REMOVE IV11??)
```{r}
##############################
# plot
##############################

MMR50_plot <- ggplot() +
  geom_point(data = epoc, aes(x = as.factor(temp), y = MO2_mmr), color = "black", alpha = 0.5, size = 3) + # width = 0.2, alpha = 0.2
    geom_point(data = IV11, aes(x = as.factor(temp), y = MO2_mmr), color = "red", alpha = 0.5, size = 3) + 
    geom_point(data = IV13, aes(x = as.factor(temp), y = MO2_mmr), color = "blue", alpha = 0.5, size = 3) +
  geom_point(data = GP07, aes(x = as.factor(temp), y = MO2_mmr), color = "green", alpha = 0.5, size = 3) +
  geom_point(data = IV16, aes(x = as.factor(temp), y = MO2_mmr), color = "purple", alpha = 0.5, size = 3) +
  geom_errorbar(data = MMR50_sum , aes(x = as.factor(temp), ymin = MMR50_avg - error, ymax = MMR50_avg + error), width = 0.1, size = 1, color = "black") +
  geom_point(data = MMR50_sum , aes(x = as.factor(temp),  y = MMR50_avg), size = 6, color = "black") +
  geom_point(data = MMR50_sum, aes(x = as.factor(temp),  y = MMR50_avg, color = as.factor(temp)), size = 4.5) +
  labs(y = expression(MMR[50] ~"(mg" ~ O[2] ~ kg^-1 ~ min^-1*")"),
       x = "Temperature (°C)") +
  scale_color_manual(values = c("lightslategray", "lightblue", "lightcoral", "indianred4")) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fill = NA, size = 0.7),
        plot.caption = element_text(size = 10, hjust = 0),
        legend.position = "none")
```

####**Part 7.** Combine plots
```{r}
##############################
# combine panels and save
##############################

epoc_plot <- cowplot::plot_grid(full_EPOC_plot, hourly_EPOC_plot, time_to_EPOC_plot, EPOC_at_MMR50_plot, time_to_MMR50_plot, MMR50_plot, nrow = 2, align = "vh", labels = "AUTO")

cowplot::save_plot(here::here("figures", "Fig?_highlights.pdf"), epoc_plot, base_width = 15, base_height = 8)
```